# Calculate connectivity
## Research question 1

> How is road connectivity affected in pre-disaster and post-disaster scenarios based on edge betweenness centrality?

### Core Metropolitan scale

@fig-cmc shows an interactive view of the connectivity before and after the flooding event in the Core Metropolitan area. On the left side, the urban arteries such as BR-290 freeway^[[https://prefeitura.poa.br/eptc/noticias/cerca-de-24-mil-veiculos-passam-pelo-corredor-humanitario-no-primeiro-dia](https://prefeitura.poa.br/eptc/noticias/cerca-de-24-mil-veiculos-passam-pelo-corredor-humanitario-no-primeiro-dia)] is displayed in purple previous to the flood event. On the right side, the flood reduced the connectivity of the urban arteries. Municipalities such as Nova Santa Rita were isolated from the core metropolitan area of Porto Alegre.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| output: false

# Load all the temp_data ending in gpkg saving you time
library(tictoc)
library(tidyverse)
library(sf)
tic()
temp_data <- list.files(path= "/home/ricardors/agile-gscience-2024-rs-flood/data/temp_data", pattern="\\.gpkg")
for (i in seq_along(temp_data)) {
    file_name <- tools::file_path_sans_ext(temp_data[i])
    assign(file_name, sf::read_sf(paste0("~/agile-gscience-2024-rs-flood/data/temp_data/", temp_data[i])))
}
# Load non-gpkg files
centrality_both_sankey_v3 <- read.csv("~/agile-gscience-2024-rs-flood/data/temp_data/centrality_both_sankey_v3.csv")
df_centrality_both_sankey <- read.csv("~/agile-gscience-2024-rs-flood/data/temp_data/df_centrality_both_sankey.csv")
# Work on flat 
sf::sf_use_s2(FALSE)
# Load gpkg files for RQ
results_data <-list.files(path= "/home/ricardors/agile-gscience-2024-rs-flood/data/results_data", pattern="\\.gpkg")
for (i in seq_along(results_data)) {
    file_name <- tools::file_path_sans_ext(results_data[i])
    assign(file_name, sf::read_sf(paste0("~/agile-gscience-2024-rs-flood/data/results_data/", results_data[i])))
} 
toc()

```


::: {#fig-cmc}

<iframe width="780" height="500" src="data/media/pre_post_cm_centrality.html" title="Quarto Documentation"></iframe>

A comparison of the centrality before and after the floods in the core metropolitan (CM) area
:::


```{r}
#| eval: false
#| echo: true
#| code-summary: "How to visualiza CM connectivity"

library(mapview)
library(leaflet)
library(leafpop)

# color: https://gis.stackexchange.com/questions/312027/r-and-mapview-how-can-i-chose-different-colors-for-different-layers

### For pre-event: 1644, 468, 142
pal <- c("#b3cde2","#8a95c5","#8854a7","#7a0f76")
centrality_pre$centrality_fct <- cut(centrality_pre$centrality,
                                     breaks=c(0,725,2114,4870,11205),
                                     labels =c("low [1-726]","intermediate [726-2114]","high [2114-4870]","urban arteries [4780-1125]"),
                                     include.lowest= TRUE,
                                     right =FALSE)
#### natural breaks:  81, 230, 582
centrality_post$centrality_fct <- cut(centrality_post$centrality,
                                      breaks=c(0,725,2114,4870,11205),
                                      labels =c("low [1-726]","intermediate [726-2114]","high [2114-4870]","urban arteries [4780-1125]"),
                                      include.lowest= TRUE,
                                      right =FALSE)
### 
library(scales)
centrality_pre$scaled_lwd <- rescale(centrality_pre$centrality, to = c(0.5, 8)) 
centrality_pre_map <- mapview::mapview(centrality_pre,
                                       zcol="centrality_fct",
                                       lwd = centrality_pre$scaled_lwd,
                                       col.regions=list("#b3cde2","#8a95c5","#8854a7","#7a0f76"),
                                       col=list("#b3cde2","#8a95c5","#8854a7","#7a0f76"),
                                       map.type=c("OpenStreetMap","CartoDB.DarkMatter"),
                                       layer.name ="Centrality Pre-Event",
                                       popup=leafpop::popupTable(centrality_pre, 
                                                        zcol=c("centrality","bidirectId"))) 
centrality_post_map <- mapview::mapview(centrality_post,
                                        zcol="centrality_fct",
                                        col.regions=list("#b3cde2","#8a95c5","#8854a7","#7a0f76"),
                                        col=list("#b3cde2","#8a95c5","#8854a7","#7a0f76"),
                                        map.type=c("OpenStreetMap","CartoDB.DarkMatter"),
                                        lwd = centrality_pre$scaled_lwd,
                                        layer.name ="Centrality Post-Event",
                                        popup=leafpop::popupTable(centrality_post, 
                                                         zcol=c("centrality","bidirectId"))) 

centrality_pre_map | centrality_post_map + mapview::mapview(flood_extent,
                                                   col.regions="#dbc3b0",
                                                   map.type=c("OpenStreetMap","CartoDB.DarkMatter"),
                                                   alpha.regions= 0.5,
                                                   layer.name="Flooding layer") 
```

We used pgrouting to calculate the centrality by counting the least-cost paths using pgr_dijkstra between origin and destinations (OD). The code [ECR 10] is based on Daniel i. Patterson post ^[[https://urbandatacyclist.com/2020/04/18/how-to-measure-centrality-among-bike-share-trips-using-pgrouting/](https://urbandatacyclist.com/2020/04/18/how-to-measure-centrality-among-bike-share-trips-using-pgrouting/)] and Matt OJO. The line 7 executed the function pgr_dijkstra from pgrouting creating a table of the shortest paths from the 300 OD contained in the table _od_300_snapped_v2_. Later, the line 18 added the spatial geometry from the road network saved in the table _ghs_aoi_net_largest_. A higher number of OD returned the error in handling memory allocation ^[[https://discourse.osgeo.org/t/handling-memory-allocation-with-pgrouting-in-centrality-analysis/110407/4](https://discourse.osgeo.org/t/handling-memory-allocation-with-pgrouting-in-centrality-analysis/110407/4)], so we limited to 300 origins and destinations. However, the visual inspection of the spatial connectivity patterns were very similar between 100 and 300 OD revealing the BR-290 freeway as the main urban arteria in both cases. 

```{sql}
#| eval: false
#| echo: true
#| code-summary: "[ACR 10] How to calculate edge betweenness centrality in the Core Metropolitan Area"

EXPLAIN ANALYZE
CREATE TABLE centrality_ghs_aoi_pre AS
SELECT
    b.id,
    b.geom,
    count(geom) as centrality
FROM pgr_dijkstra(
    'SELECT 
        id AS id,
        "fromId" AS source,
        "toId" AS target,
        weight as cost
    FROM 
      ghs_aoi_net_largest',
    ARRAY(SELECT snapped_id AS start_id FROM od_300_snapped_v2),
    ARRAY(SELECT snapped_id AS end_id FROM od_300_snapped_v2),
        directed:=TRUE) j
    LEFT JOIN 
      ghs_aoi_net_largest AS b
    ON
      j.edge=b.id
    GROUP BY
        b.id, b.geom;        
		
--- Time: 80542.531 ms

```

The code chunks from ACR11 to ACR13 handles a characteristic from exporting the ORS road network into R, the bidirectid variable. 

```{sql}
#| eval: false
#| echo: true
#| code-summary: "[ACR 11] How to add a bidirectId when not present"

CREATE TABLE centrality_ghs_aoi_bidirect_pre AS
SELECT 
	t1.*,
	t2."bdrctId"
FROM
	centrality_ghs_aoi_pre AS t1
JOIN
	ghs_aoi_net_largest AS t2
ON
t1.id = t2.id;
--- Time: 164.685 ms

select max("bdrctId") from centrality_ghs_aoi_bidirect_pre ; ---99527
create sequence bidirectid_centrality_ghs_aoi start 99527;
update centrality_ghs_aoi_bidirect_pre
set "bdrctId" = nextval('bidirectid_centrality_ghs_aoi')
where "bdrctId" is null ;
--- checking that bdrctId is not null
select count(*)
from centrality_ghs_aoi_bidirect_pre
where "bdrctId" is null; ---0
```

We summed up the centrality values of the roads with same bidirectid.  

```{sql}
#| eval: false
#| echo: true
#| code-summary: "[ACR 12] How to use bidirectid to calculcalte centrality"

create table centrality_ghs_aoi_bidirect_pre_group as
select
"bdrctId",
sum(centrality) as centrality
from centrality_ghs_aoi_bidirect_pre
group by "bdrctId";
--- Time: 65.993 ms
create table centrality_ghs_aoi_bidirect_pre_group_id as
select t1.*, t2.id
from centrality_ghs_aoi_bidirect_pre_group t1
join centrality_ghs_aoi_bidirect_pre t2
on t1."bdrctId" = t2."bdrctId";
```


```{sql}
#| eval: false
#| echo: true
#| code-summary: "[ACR 13] How to calculate final edge betweenness centrality in the Core Metropolitan Area"

create table centrality_ghs_aoi_bidirect_pre_cleaned as
select
	t1.*,
	t2.geom,
	t2."toId",
	t2."fromId",
	t2.weight,
	t2."undrctI"
from
	centrality_ghs_aoi_bidirect_pre_group_id t1
join
	ghs_aoi_net_largest t2
on
t1.id = t2.id;
--- Time: 565.951 ms
```

This process is repeated again changing the table containing the road network _ghs_aoi_net_largest_  to _ghs_aoi_net_largest_post_. While the table _ghs_aoi_net_largest_  represents the road network of the AoI before the flood, the table _ghs_aoi_net_largest_post_ shows the differences on the road network after the flood.

### Intracity scale

@fig-lc shows the changes on the centrality when we limited the AoI to each municipality, in this case the municipality of Canoas. The flood hit the west side of Canoas affecting the Mathias Velho dam ^[[https://prefeitura.poa.br/eptc/noticias/cerca-de-24-mil-
veiculos-passam-pelo-corredor-humanitario-no-primeiro-dia](https://prefeitura.poa.br/eptc/noticias/cerca-de-24-mil-
veiculos-passam-pelo-corredor-humanitario-no-primeiro-dia)] causing the loss of urban arterias such as Rua Florianópolis.

::: {#fig-lc}

<iframe width="780" height="500" src="data/media/pre_post_ic_centrality.html" title="Quarto Documentation"></iframe>

A comparison of the centrality before and after the floods in the core metropolitan area
:::


```{r}
#| echo: true
#| eval: false
#| message: false
#| warning: false
#| code-summary:  "How to visualiza a comparison of the centrality before and after the floods in the intracity (IC) area"


library(leaflet)
library(mapview)
library(leafpop)

### For pre-event: 1644, 468, 142
pal <- c("#b3cde2","#8a95c5","#8854a7","#7a0f76")
local_connectivity_canoas <- filter(local_connectivity_pre, shapename=="Canoas")
local_connectivity_canoas_post <- filter(local_connectivity_post, shapename=="Canoas")
library(BAMMtools)
natural_jenks_breaks_cmc_pre_canoas <- BAMMtools::getJenksBreaks(local_connectivity_canoas$centrality, k=5)

local_connectivity_canoas$centrality_fct <- cut(local_connectivity_canoas$centrality,
                                     breaks=c(
                                              natural_jenks_breaks_cmc_pre_canoas[1],
                                              natural_jenks_breaks_cmc_pre_canoas[2],
                                              natural_jenks_breaks_cmc_pre_canoas[3],
                                              natural_jenks_breaks_cmc_pre_canoas[4],
                                              natural_jenks_breaks_cmc_pre_canoas[5]),
                                     labels =c("low [1-533]","intermediate [533-1510]","high [1510-2899]","urban arteries [2899-5727]"),
                                     include.lowest= TRUE,
                                     right =FALSE)
#### natural breaks:  81, 230, 582
local_connectivity_canoas_post$centrality_fct <- cut(local_connectivity_canoas_post$centrality,
                                      breaks=c(
                                              natural_jenks_breaks_cmc_pre_canoas[1],
                                              natural_jenks_breaks_cmc_pre_canoas[2],
                                              natural_jenks_breaks_cmc_pre_canoas[3],
                                              natural_jenks_breaks_cmc_pre_canoas[4],
                                              natural_jenks_breaks_cmc_pre_canoas[5]),
                                      labels =c("low [1-533]","intermediate [533-1510]","high [1510-2899]","urban arteries [2899-5727]"),
                                      include.lowest= TRUE,
                                      right =FALSE)
### 
library(scales)
local_connectivity_canoas$scaled_lwd <- rescale(local_connectivity_canoas$centrality, to = c(0.5, 8)) 
local_connectivity_canoas_pre_map <- mapview::mapview(local_connectivity_canoas,
                                       zcol="centrality_fct",
                                       lwd = local_connectivity_canoas$scaled_lwd,
                                       col.regions=list("#b3cde2","#8a95c5","#8854a7","#7a0f76"),
                                       col=list("#b3cde2","#8a95c5","#8854a7","#7a0f76"),
                                       map.types=c("OpenStreetMap","CartoDB.DarkMatter"),
                                       layer.name ="IC connectivity Pre-Event",
                                       popup=leafpop::popupTable(local_connectivity_canoas, 
                                                        zcol=c("centrality","bidirectid"))) 
local_connectivity_canoas_post_map <- mapview::mapview(local_connectivity_canoas_post,
                                       zcol="centrality_fct",
                                       lwd = local_connectivity_canoas$scaled_lwd,
                                       col.regions=list("#b3cde2","#8a95c5","#8854a7","#7a0f76"),
                                       col=list("#b3cde2","#8a95c5","#8854a7","#7a0f76"),
                                       map.types=c("OpenStreetMap","CartoDB.DarkMatter"),
                                       layer.name ="IC connectivity Post-Event",
                                       popup=leafpop::popupTable(local_connectivity_canoas_post, 
                                                        zcol=c("centrality","bidirectid"))) 

local_connectivity_canoas_pre_map | local_connectivity_canoas_post_map + mapview::mapview(flood_extent,
                                                   col.regions="#dbc3b0",
                                                   map.types=c("OpenStreetMap","CartoDB.DarkMatter"),
                                                   alpha.regions= 0.5,
                                                   layer.name="Flooding layer")
```


The code chunk [ECR14] shows how to calculate the IntraCity (IC) connectivity combining SQL queries with the R library glue [@glue]. We used the same central query in line 20 that calculated the least-cost paths. However, we created a routable network for each municipality with the iteration between the line 7 and 40. 

```{r}
#| eval: false
#| echo: true
#| code-summary: "[ECR 14] How to calculate edge betweenness centrality in the Core Metropolitan Area"

## Table with networks
library(sf)
table_rq_1_2_net <- st_read(connection, "table_rq_1_2_net")
table_snapped <- st_read(connection, "table_snapped")
municipalities_ghs <- st_read(connection, "municipalities_ghs")
  
tic()  
for(idx in 1:9){
  tic(glue("run {idx}/9")) 
  table_name_1_centrality <- glue("net_largest_{as.character(idx)}")
  temp_table_name <- glue("temp_centrality_{idx}")
  table_bidirect_id <- glue("t_bidirect_id_{idx}")
  dbExecute(glue_sql("DROP TABLE IF EXISTS {`temp_table_name`}", .con=connection),
          conn=connection)  
query_1_central <- glue_sql("CREATE TABLE {`temp_table_name`} AS
 SELECT   
 b.id,
 b.geom,
 count(b.geom) AS centrality 
 FROM  pgr_dijkstra('SELECT id,
                            fromid AS source,
                            toid AS target,
                            weight AS cost
                      FROM {`table_name_1_centrality`}',
                      ARRAY(SELECT 
                        snapped_id AS start_id 
                        FROM table_snapped 
                        WHERE shapename_id = {idx}::text),
                      ARRAY(SELECT
                        snapped_id AS end_id 
                        FROM table_snapped
                        WHERE shapename_id = {idx}::text),
                      directed := TRUE) j
                      LEFT JOIN {`table_name_1_centrality`} AS b
                      ON j.edge = b.id
                      GROUP BY  b.id, geom;", .con=connection)

dbExecute(con=connection, query_1_central)
toc()
}
toc()
# Time: 53805 ms
```


```{sql}
#| eval: false
#| echo: false

---- 3 
 CREATE TABLE centrality_bidirect_temp_centrality_3 AS
SELECT t1.*,
    t2."bidirectid"
FROM
    temp_centrality_3 t1
JOIN
    net_largest_3 t2
ON
    t1.fid = t2.fid;
--- The final product requries  79941   
select max("bidirectid") from centrality_bidirect_temp_centrality_3 ;
create sequence bididirect_id_temp_centrality_3_seq start 99494;
update centrality_bidirect_temp_centrality_3
set "bidirectid" = nextval('bididirect_id_temp_centrality_3_seq')
where "bidirectid" is null ;
---- verify
select count(*) 
from centrality_bidirect_temp_centrality_3 
where "bidirectid" is null; --- 0
----
create table centrality_bidirect_temp_centrality_3_bidirect_group as 
select 
       "bidirectid",
       sum(centrality) as  centrality
from centrality_bidirect_temp_centrality_3 
group by "bidirectid"; --- this sum the centrality for duplicated bidirect id
---
---- now recover the id
create table centrality_bidirect_temp_centrality_3_bidirect_group_id as 
select t1.*, t2.fid
from centrality_bidirect_temp_centrality_3_bidirect_group t1
join centrality_bidirect_temp_centrality_3 t2 
on t1."bidirectid" = t2."bidirectid"; 
---- add geometries
create table centrality_bidirect_temp_centrality_3_cleamed as
select t1.*,
    t2.the_geom,
    t2.toid,
    t2.fromid,
    t2.weight,
    t2."unidirectid"
from 
    centrality_bidirect_temp_centrality_3_bidirect_group_id t1
join
    net_largest_3 t2
on
  t1.fid = t2.fid;
```

## Research question 2

> Which municipalities were most affected by the flood based on the edge betweenness centrality metrics?

The @tbl-municipalities aggregate the centrality values for each municipalities. Nova Santa Rita is the most affected municipality when the CM centrality is assessed. However, this same municipality ranked as 6th in terms of loss of IC connectivity. 

```{r}
#| code-summary: "How to create an interactive table that show municipalities loss of centrality"
#| echo: true
#| message: false
#| waning: false
#| label: tbl-municipalities
#| tbl-cap: "Pop.Aff: Population affected. CMC: Core Metropolitan Connectivity. IC: IntraCity connectivity. GDP: Gross domestic product per capita (2021)"

library(DT)
 sf::st_drop_geometry(connectivity_municipality_lc_rq) |>
  left_join(sf::st_drop_geometry(connectivity_municipality_cmc_rq),
            by=join_by(gid)) |> 
  mutate(diff_ic = (round((post_lc -pre_lc)/pre_lc*100)),
         diff_cmc = (round((post_gc - pre_gc)/pre_gc*100)),
         pop_aff = paste0(pop_aff.x/1000,"K (", round(pop_aff.x/pop.x*100),"%)"),
         gdp = paste0(pib_mun.x/1000,"K")) |>
  select(municipality.x, pop_aff, diff_cmc, diff_ic,gdp) |> 
DT::datatable(colnames=c("Municipality","Pop.Aff","CMC","IC","GDP(R$"),
              extensions = 'Buttons',
              filter = "top",
              options=list(
                          dom = 'Bfrtip',
                          columnDefs = list(list(className = 'dt-center', targets = c(1:4))),
                          pageLength = 5,
                           buttons = c('copy', 'csv')),
              escape = FALSE) 
```

The @fig-mun 

```{r}
#| echo: true
#| warning: false
#| message: false
#| eval: false
#| code-summary: "How to visualise a comparison of the centrality before and after the floods by municipalities"

library(scales)
library(RColorBrewer)
library(mapview)
# Define a red-to-green palette
palette_red_green <- colorRampPalette(c("red", "yellow", "green"))

connectivity_municipality_cmc_rq$diff_cmc <-  round(((connectivity_municipality_cmc_rq$post_gc - connectivity_municipality_cmc_rq$pre_gc) / connectivity_municipality_cmc_rq$pre_gc*100))
connectivity_municipality_lc_rq$diff_lc <-  round(((connectivity_municipality_lc_rq$post_lc - connectivity_municipality_lc_rq$pre_lc) / connectivity_municipality_lc_rq$pre_lc*100))

municipalities_pre_map <- mapview::mapview(connectivity_municipality_cmc_rq,
                                       zcol="diff_cmc",
                                       map.types=c("OpenStreetMap","CartoDB.DarkMatter"),
                                       layer.name ="Core Metropoplitan Connectivity (CM)",
                                       col.regions = palette_red_green(6),
                                       popup=leafpop::popupTable(connectivity_municipality_cmc_rq, 
                                                        zcol=c("municipality","diff","pop_aff","pib_mun"))) 
municipalities_post_map <-  mapview::mapview(connectivity_municipality_lc_rq,
                                       zcol="diff_lc",
                                       map.types=c("OpenStreetMap","CartoDB.DarkMatter"),
                                       layer.name ="IntraCity Connectivity (IC)",
                                       col.regions = palette_red_green(6),
                                       hide=TRUE,
                                       popup=leafpop::popupTable(connectivity_municipality_lc_rq, 
                                                        zcol=c("municipality","diff","pop_aff","pib_mun"))) 

municipalities_pre_map + (municipalities_post_map + mapview(flood_extent, hide =TRUE, col.regions="#dbc3b0", alpha.regions=.5) )

```


::: {#fig-mun}

<iframe width="780" height="500" src="data/media/pre_post_municipality_centrality.html" title="Quarto Documentation"></iframe>

A comparison of the centrality before and after the floods aggregated by municipalities
:::
