# Select AoI

The main objects of the research were the administrative boundaries from the municipalities, the flood extent, road network and hospitals. 

* <p> geoBoundaries provided the <span class="font-color">**administrative units (ADM2)**</span> of Brazil [@Runfola2020]. We subset the data using 9 municipalities defined as the core metropolitan area of Rio Grande do Sul. The reason why we selected these municipalities is that they intersected with the dense urban center 1210 from the Global Human Settlement Layer [@https://doi.org/10.2905/a0df7a6f-49de-46ea-9bde-563437a6e2ba].</p>
* <p> The Federal University of Rio Grande do Sul published the <span class="font-color">**flood extent**</span> observed from the Skystat, Planet and WorldView-2 satellites and also validated through field surveys [@https://doi.org/10.5281/zenodo.13999016].</p>
* <p> The <span class="font-color">**road network**</span> is downloaded from GeoFabrik^[[https://download.geofabrik.de/south-america/brazil/sul.html#](https://download.geofabrik.de/south-america/brazil/sul.html#)]. </p>
* The State Health Department shared the capacity of Intense Care Units (ICU) beds [@https://doi.org/10.7303/syn32211006.1], while the "Secretaria Estadual da Saúde/DGTI" published the <span class="font-color">**healthcare facilities**</span> in Rio Grande do Sul with ICU beds using a [WFS](https://iede.rs.gov.br/server/rest/services/SES/Hospitais_Leitos_UTI_RS/FeatureServer).

```{=html}
<iframe width="780" height="500" src="data/media/import_data_v0.html" title="Quarto Documentation"></iframe>
```


```{r}
#| echo: true
#| code-summary: "R code to create interactive map based on leaflet"
#| eval: false

library(leaflet.extras)
library(leaflet)
leaflet() |>
  addTiles()  |>
  addProviderTiles("OpenStreetMap", group = "OpenStreetMap")  |>
  addProviderTiles("Esri.WorldImagery", group = "Esri.WorldImagery")  |>
  addPolygons(data=municipalities_ghs_leaflet,
                      weight = 2,
                      color = "black",
                      fillOpacity = 0,
                      dashArray = "3",
                      popup = ~paste0( "<b>Municipality: </b>", shapename, "<br/>",
                                      "<b>Population affected: </b>", Pop.Aff, "<br/>",
                                      "<b>PIB per capita: </b>", GDP, "<br/>"),
                                      group= "Municipalities (CMA)") |>
  addPolygons(data=urban_center_4326_core,
                    fillColor = "yellow",
                    opacity = 0.2,
                    weight = 2,
                    color = "black",
                    dashArray = "1",
                    group= "Dense Urban Center Settlment",
                    popup = ~paste0( "<b>Settlment ID: </b>", gid, "<br/>",
                    "<b>Population: </b>", round(pop_2020), "<br/>",
                    "<b>Building Surface (m2): </b>", round(bu_m2_2020))) |>
                    addPolygons(data=flooding_simplified_porto_united,
                    fillColor = "#dec8b7ff",
                    fillOpacity = 0.6,
                    weight = 1,
                    color = "#4b2609",
                    dashArray = "1",
                    group= "Flood extent") |>
  addCircles(data=filter(hospital_bed_aux_leaflet, nat_jur_cat=="Public"),
                    radius=10,
                    weight=15,
                    opacity = 0.6,
                    color = "#6da991ff",
                    popup= ~paste0("<b>Hospital: </b>", stringr::str_to_title(ds_cnes), "<br/>",
                    "<b>Type Jur: </b>", nat_jur_cat, "<br/>",
                    "<b>Nature Jur: </b>", stringr::str_to_title(nm_razao_s), "<br/>",
                    "<b>UCI Beds: </b>", beds,"<br/>",
                    "<b>Risk: </b>", tp_risco),
                    group="Hospital (public)")  |>
  addCircles(data=filter(hospital_bed_aux_leaflet, nat_jur_cat=="Private"),
                    radius=10,
                    weight=15,
                    opacity = 0.6,
                    color = "#3f98c3ff",
                    popup= ~paste0("<b>Hospital: </b>", stringr::str_to_title(ds_cnes), "<br/>",
                    "<b>Type Jur: </b>", nat_jur_cat, "<br/>",
                    "<b>Nature Jur: </b>", stringr::str_to_title(nm_razao_s), "<br/>",
                    "<b>UCI Beds: </b>", beds,"<br/>",
                    "<b>Risk: </b>", tp_risco),
                    group="Hospital (private)")   |>
  addRasterImage(building_density_local_aoi_4326_masked,
                    opacity = 0.5,
                    group="Building density") |>
                    addLayersControl(
                    baseGroups = c("OpenStreetMap",
                    "Esri.WorldImagery"),
                    overlayGroups = c("Municipalities (CMA)",
                    "Hospital (public)",
                    "Hospital (private)",
                    "Dense Urban Center Settlment",
                    "Flood extent",
                    "Building density"),
                    options = layersControlOptions(collapsed = FALSE)) |>
                    hideGroup(c("Dense Urban Center Settlment",
                    "Building density"))
```


## Importing data into PostgreSQL

```{bash}
#| eval: false
#| echo: true
#| code-summary: "Importing source data using ogr2ogr and shp2pgsql"
# Urban Center to select core metropolitan area (2.514s)
time shp2pgsql -D -I -s 54009:4326 GHS_SMOD_E2020_GLOBE_R2023A_54009_1000_UC_V2_0.shp agile_gis_2025_rs.urban_center_4326 | psql -p 25432 -U docker -d gis  -h localhost
# Hospitals (0.217s)
time ogr2ogr -f PostgreSQL PG:"host=localhost port= 25432 user=docker password=docker dbname=gis schemas=agile_gis_2025_rs" Hospitais_com_Leitos_de_UTIs_no_RS.geojson -nln hospitals 
# Administrative units (5.854s)
time ogr2ogr -f PostgreSQL PG:"host=localhost port= 25432 user=docker password=docker dbname=gis schemas=agile_gis_2025_rs" geoBoundaries-BRA-ADM2.geojson -nln nuts
#  Flood extent (1.008s)
time ogr2ogr -f PostgreSQL PG:"host=localhost port= 25432 user=docker password=docker dbname=gis schemas=agile_gis_2025_rs" rhguaiba_planetskysat_inundacao_obs_20240506.gpkg -nln flooding_raw
```

## Connecting R with PostgreSQL

```{r}
#| echo: true
#| eval: false
library(DBI)
library(dplyr)
## connect to PostgreSQL
connection <- DBI::dbConnect(RPostgres::Postgres(),
                             user="docker",
                             password="docker",
                             host="localhost",
                             dbname="gis",
                             port=25432)
## Connect to PostgreSQL to read CSV without defining in PostgreSQL
library(tictoc)
tic()
etlcnes_hospital_selected <- read.csv("~/agile_giscience_2025_github/agile-gscience-2024-rs-flood/data/source_data/ETLCNES_SR_RS_21_12_t.csv") |> 
                          dplyr::select(c("CNES","QTLEITP1","QTLEITP2","QTLEITP3","NAT_JUR"))
toc() # 54.361 sec elapsed
## Write table using PostgreSQL connection
DBI::dbWriteTable(connection, DBI::Id(schema="agile_gis_2025_rs",table="etlcnes_hospital_selected"))
```


```{r}
#| echo: false
#| eval: false

hospital_bed_aux_leaflet <- sf::st_read(connection, DBI::Id(schema="agile_gis_2025_rs" ,"hospital_bed_aux_leaflet"))
municipalities_ghs_leaflet <- sf::st_read(connection, DBI::Id(schema="agile_gis_2025_rs" ,"municipalities_ghs_leaflet"))
hospital_bed_aux_leaflet <- sf::st_read(connection, DBI::Id(schema="agile_gis_2025_rs" ,"hospital_bed_aux_leaflet"))
hospital_bed_aux_leaflet <- sf::st_read(connection, DBI::Id(schema="agile_gis_2025_rs" ,"hospital_bed_aux_leaflet"))
flooding_simplified_porto_united <- sf::st_read(connection, DBI::Id(schema="agile_gis_2025_rs" ,"flooding_simplified_porto_united"))
urban_center_4326_core <- sf::st_read(connection, DBI::Id(schema="agile_gis_2025_rs" ,"urban_center_4326_core"))
municipalities_ghs <- sf::st_read(connection, DBI::Id(schema="agile_gis_2025_rs" ,"municipalities_ghs"))
```

## Data wranging

### Vector data

```{sql}
#| eval: false
#| echo: true

---- Dense urban center that intersects with Porto Alegre -> ***Input for leaflet map***
EXPLAIN ANALYZE
CREATE TABLE urban_center_4326_core AS
SELECT 
	duc.* 
FROM 
	urban_center_4326 AS duc,
	(SELECT 
		* 
	 FROM 
	 	nuts_shp 
	 WHERE
	 	shapename = 'Porto Alegre') AS porto_alegre
WHERE
	st_intersects(
		duc.geom,
		porto_alegre.geom);
--- 13.714 ms

--- Municipalities contained in the urban_center_4326
EXPLAIN ANALYZE	
CREATE TABLE municipalities_ghs AS
WITH porto_alegre_ghs AS(
    SELECT 
       ghs.*
    FROM
       urban_center_4326 AS ghs
    JOIN
       nuts 
    ON 
       st_intersects(nuts.wkb_geometry, ghs.geom)
    WHERE 
        nuts.shapename = 'Porto Alegre'),
municipalities_ghs_geom AS (
	SELECT
		municipalities.*
	FROM 
		nuts AS municipalities,
		porto_alegre_ghs AS ghs
	WHERE
		st_intersects(municipalities.wkb_geometry, ghs.geom))
	SELECT * FROM municipalities_ghs_geom; --- 19.062ms
	
---- Defined core metropolitan area of the municipalities_ghs
EXPLAIN ANALYZE
CREATE TABLE core_metropolitan_area AS
SELECT 
	st_union(wkb_geometry) AS geom
FROM 
municipalities_ghs; --- 15.382ms
---- Hospitals with ICU and their bed capacity -> ***Input for leaflet map****


EXPLAIN ANALYZE
CREATE TABLE hospital_bed_aux_leaflet AS
SELECT 
	h.fid,
	initcap(h.ds_cnes) AS ds_cnes,
	h.tp_risco,
	h.nm_municip,
	h.ds_bairro,
	h.addr_ct,
	h.geom_hospital,
	h.distance,
		aux_data."QTLEITP1",
	aux_data."QTLEITP2",
	aux_data."QTLEITP3",
	"QTLEITP1"+"QTLEITP2"+"QTLEITP3" AS beds,
	aux_data."NAT_JUR",
	CASE
		WHEN "NAT_JUR" IN (3069,3999)
		THEN 'Private' ELSE 'Public'  
	END AS nat_jur_cat,
	initcap(h.nm_razao_s) AS nm_razao_s
FROM 
	etlcnes_hospital_selected AS aux_data
JOIN
	municipalities_ghs_hospitals AS h
ON
	h.cd_cnes = aux_data."CNES"; ---- Execution Time: 43.901 ms
	
--- Municipality of the AoI with additonial data from:
------- A) IBGR: "https://www.ibge.gov.br/cidades-e-estados/rs/"
--------B) DEE-SPGG: "https://mup.rs.gov.br/" 
--- -> ***Input for leaflet map****
EXPLAIN ANALYZE
CREATE TABLE municipalities_ghs_leaflet AS
SELECT shapename,
	 wkb_geometry
FROM municipalities_ghs; ----  8.511 ms
--- Add new fields (A,B)
ALTER TABLE municipalities_ghs_leaflet
	ADD COLUMN "Pop.Aff" text,
	ADD COLUMN "GDP" text ; --- 0.012S execute time
--- Populate columns
EXPLAIN ANALYZE
	UPDATE municipalities_ghs_leaflet
SET "Pop.Aff" = data."Pop.Aff",
    "GDP" = data."GDP"
FROM (VALUES
    ('Alvorada', '26K (29%)', '15K'),
    ('Cachoeirinha', '12K (9%)', '49K'),
    ('Canoas', '157K (45%)', '63K'),
    ('Esteio', '20K (26%)', '45K'),
    ('Gravataí', '6K (2%)', '36K'),
    ('Nova Santa Rita', '7K (24%)', '81K'),
    ('Porto Alegre', '125K (9%)', '55K'),
    ('Sapucaia do Sul', '6K (4%)', '29K'),
    ('Viamão', '2K (1%)', '17K')
) AS data(shapename, "Pop.Aff", "GDP")
WHERE municipalities_ghs_leaflet.shapename = data.shapename; --- Execution Time: 0.596 ms

```


### Raster data

We used a subset of the GHS_BUILT_V available at origin and zenodo repository 

```{r}
#| echo: true
#| eval: false

# Import data
library(sf)  # vector files
library(terra) # raster files
library(tictoc) # timing
tic()
buildings_density_global <- terra::rast(
  '~/GHS_BUILT_V_E2020_GLOBE_R2023A_54009_100_V1_0.tif')
toc() ## 0.088 sec elapsed
tic()
municipalities_ghs_transformed <- municipalities_ghs |> st_transform(crs(buildings_density_global))
toc() ## 0.071 sec elapsed
tic()
building_density_local_aoi <- crop(buildings_density_global, municipalities_ghs_transformed)
toc() ## 0.074 sec elapsed
## Reproject for sampling
tic()
building_density_local_aoi_4326 <- terra::project(building_density_local_aoi, crs(municipalities_ghs))
toc() ### 0.595 sec elapsed
tic()
building_density_local_aoi_4326_masked <- terra::mask(building_density_local_aoi_4326, municipalities_ghs) 
toc() ## 0.089 sec elapsed
terra::writeRaster(building_density_local_aoi_4326_masked, "GHS_BUILT_V_E2020_GLOBE_R2023A_4326_100_V1_0_RioGrandeDoSul.tif")
### Reclassify values 0 with missing values (NA) to make them transparent in the leaflet map
values(building_density_local_aoi_4326_masked)[values(building_density_local_aoi_4326_masked) == 0] = NA
```

