# Origin-Destination matrix

```{r}
#| eval: true
#| echo: false
#| message: false

library(DBI)
library(dplyr)
library(RPostgres)
## connect to PostgreSQL
connection <- DBI::dbConnect(RPostgres::Postgres(),
                             user="docker",
                             password="docker",
                             host="localhost",
                             dbname="gis",
                             port=25432)
od_300_v2 <- sf::st_read(connection, Id(schema="agile_gis_2025_rs",table="od_300_v2"))
od_200_each <- sf::st_read(connection, Id(schema="agile_gis_2025_rs",table="od_200_each"))
od_300_snapped_v2 <- sf::st_read(connection, Id(schema="agile_gis_2025_rs",table="od_300_snapped_v2"))
```


We created Origin and Destination (OD) points to create shortest path in each of the combinations using the costs from ORS instead of the distance. 

For the centrality, we used the Global Human Settlement Built-up volume raster to obtain a <span class='font-color'>**weighted OD**</span>. Those regions with a higher built-up volume had a higher number of OD. The following map shows the weighted OD based on the building density that were used to calculate the core metropolitan centrality of the AoI (Core Metropolitan Area) and to calculate the intracity connectivity (Municipality).

```{=html}
<iframe width= "780" height="500" src="data/media/weighted_od.html" title "Weighted OD"></iframe>
```


```{r}
#| eval: false
#| echo: true
#| code-summary: "How to create the map that shows weighted OD sampling to calculate centrality of the road network"

library(sf)
library(DBI)

## Import Data from PostgreSQL database
municipalities_ghs <- st_read(connection,
        DBI::Id(schema="agile_gis_2025_rs",
                table="municipalities_ghs"))
core_metropolitan_area <- st_read(connection,
        DBI::Id(schema="agile_gis_2025_rs",
                table="core_metropolitan_area"))        
od_300_snapped_v2 <- st_read(
                              connection,
                              DBI::Id(schema="agile_gis_2025_rs",
                                      table="od_300_snapped_v2"))
## Create interactive map
### styles
library(raster)
pal <- colorNumeric(c("#faebdd", "#cd1c4e", "#110c24"), values(build_masked),
  na.color = "transparent")
pal_fct <- colorFactor(
  palette = c('#e41a1c', '#377eb8', '#4daf4a', '#984ea3', '#ff7f00','#ffff33','#a65628', '#f781bf','#999999' ),
  domain = od_200_each$shapename
)
### map
library(leaflet)
leaflet() |>
  addTiles()  |>
  addProviderTiles("CartoDB.Positron", group = "CartoDB.Positron")  |>
  addProviderTiles("Esri.WorldImagery", group = "Esri.WorldImagery")  |>                    addPolygons(data=municipalities_ghs,
                    fillColor = "yellow",
                    fillOpacity = 0.05,
                    weight = 1.5,
                    color = "red",
                    dashArray = "1",
                    group= "Municipalities") |> 
  addPolygons(data=core_metropolitan_area,
                    fillColor = "yellow",
                    fillOpacity = 0.05,
                    weight = 2,
                    color = "red",
                    dashArray = "1",
                    group= "Core Metropolitan Area") |> 
  addCircleMarkers(data=od_300_snapped_v2,
                    fillColor = "yellow",
                    fillOpacity = 0.8,
                    weight = 1.5,
                    radius= 3,
                    color = "black",
                    dashArray = "1",
                    group= "Weighted OD - C.Metropolitan Area") |> 
   addCircleMarkers(data=od_200_each,
                    color = ~pal_fct(shapename),
                    fillOpacity = 0.8,
                    weight = 1.5,
                    radius= 3,
                    dashArray = "1",
                    group= "Weighted OD - Municipality") |> 
  addRasterImage(build_masked,
                 opacity=0.7,
                 colors = pal,
                 group="Bulding density") |> 
     addLegend(pal = pal, values = values(build_masked),
    title = "Building Density (m³)") |> 
 addLayersControl(
    baseGroups = c("CartoDB.Positron", "Esri.WorldImagery"),
    overlayGroups = c("Municipalities","Core Metropolitan Area", "Weighted OD - C.Metropolitan Area","Weighted OD - Municipality","Bulding density"),
    options = layersControlOptions(collapsed = FALSE)
  ) |>
  hideGroup(c("Municipalities","Core Metropolitan Area","Weighted OD - Municipality")) |>
  setView( lng =-51.21566 , lat = -30.04176, zoom = 12 ) 
```

For the the redundancy, a component of the resiliene, we used a <span class='font-color'>**regular OD**</span> sampling in isochrones of 10 minutes for the hospitals equiped with ICU beds in the Area of Interest. The following map shows these Isochornes, hospitals as Point of Interest (PoI) and the regular OD samples.

```{=html}
<iframe width="780" height=500 src="data/media/regular_od.html" title="Regular OD"></iframe>
```


```{r}
#| echo: true
#| eval: false
#| code-summary: "How to create the map that shows regular OD sampling to calculate redundancy of the hospitals"

## Import data:
regular_sampling_isochrones_snapped_ghs_aoi <- 
  sf::st_read(connection, Id(schema="agile_gis_2025_rs",table="regular_sampling_isochrones_snapped_ghs_aoi"))

## style:
factpal <- colorFactor(topo.colors(22), isochrones_poi_hospitals$ds_cnes)
## Map
leaflet() |>
  addTiles()  |>
  addProviderTiles("CartoDB.Positron", group = "CartoDB.Positron")  |>
  addProviderTiles("Esri.WorldImagery", group = "Esri.WorldImagery")  |>   
  addProviderTiles("CartoDB.DarkMatter", group = "CartoDB.DarkMatter") |> 
  addCircleMarkers(data=hospitals_poi,
                    fillColor = "yellow",
                    fillOpacity = 0.6,
                    weight = 3,
                    radius = 5,
                    color = "red",
                    dashArray = "1",
                    group= "Hospitals PoI") |> 
  addPolygons(data=isochrones_poi_hospitals,
                    fillOpacity = 0.2,
                    weight = 2,
                    color = ~factpal(ds_cnes),
                    dashArray = "1",
                    group= "Isochrones") |> 
  addCircleMarkers(data=regular_sampling_isochrones_snapped_ghs_aoi,
                    fillOpacity = 0.8,
                    weight = 1.5,
                    radius= 3,
                    color =~factpal(ds_cnes),
                    dashArray = "1",
                    group= "Regular OD") |> 
 addLayersControl(
    baseGroups = c("CartoDB.Positron", "Esri.WorldImagery","CartoDB.DarkMatter"),
    overlayGroups = c("Hospitals PoI","Isochrones", "Regular OD"),
    options = layersControlOptions(collapsed = FALSE)
  ) |> 
    setView( lng =-51.01134 , lat = -29.93259, zoom = 11.5 ) 
```

**Troubleshootin 10** There were several tables when exporting _od_300_v2_ to GIS from PostgreSQL database. The reason is that unlike PostgreSQL, QGIS only admit one geometry type column. We solved this checking the attribute table in QGIS and choosing the table accordingly. 

```{r}
#| eval: false
#| include: false
#| echo: false


# Snapping
od_200_each <- st_read(connection, Id(schema="agile_gis_2025_rs", table="od_200_each" ))
od_200_each_group_id <- st_read(connection, Id(schema="public", table="od_200_each_group_id"))  |> 
                          mutate(dense_rank = as.character(dense_rank)) 
```


## Weighted OD

### Create weighted sampling points
#### For the Core Metropolitan Area

We masked the global dataset GHS-Built-up volume to the area of interest named Core Metropolitan Area and took 300 samples based the building density. A higher number of points caused the `ERROR:  invalid memory alloc request size 2450771848`, however, a possible solution is to iterate using batch queries. Although the limitation of the OD to 300 points could be an underrepresented of the phenomena, we did observed similar patterns in the centrality with visual inspections. The function spatSample from the library terra [@terra] created the table _od_300_v2_ that is used to calculate the core metropolitan centrality.

```{r}
#| eval: false
#| echo: true
#| code-summary: "[ACR-1] How to take weighted OD samples in the Core Metropolitan Area"

# Import data
library(sf) # handle vector files
library(terra) # handle raster files
buildings <- terra::rast("/home/ricardors/agile-gscience-2024-rs-flood/data/source_data/GHS_BUILT_V_E2020_GLOBE_R2023A_4326_100_V1_0_RioGrandeDoSul.tif")
pop_ghs <- 300 # set up number of sampling points
## Select area of interest from the database
municipalities_ghs <- sf::read_sf(connection,
                                  DBI::Id(schema="agile_gis_2025_rs",
                                          table="municipalities_ghs"))
municipalities_ghs_reproj <- municipalities_ghs |>
                                            st_transform(crs(buildings))
build_cropped <- crop(buildings,
                      municipalities_ghs_reproj)
## Reproject for sampling
build_4326 <- terra::project(build_cropped,
                             crs(municipalities_ghs))
build_masked <- terra::mask(build_4326,
                            municipalities_ghs)
## Weighted sampling
set.seed(123)
od <- spatSample(build_masked,
                     pop_ghs,
                    "weights",
                     as.points = TRUE) |>
                st_as_sf() |>
                st_transform(4326)
## Rename column
names(od)[1] <- "building"
## Upload the sampling points to the database
sf::dbWriteTable(connection, DBI::Id(schema = "agile_gis_2025_rs",
table = "od_300_v2"), od)

# time: 721 ms
```

```{r}
#| echo: false
#| message: false
library(kableExtra)
kbl(slice(od_300_v2,1:5))
```

#### For the Intracity connectivity

We iterated through each municipality masking the GHS-Built-up volume and taking 200 weighted OD samples. The list that stored all the iterations is converted into a spatial dataframe (sf).

```{r}
#| eval: false
#| echo: true
#| code-summary: "[ACR-2] How to take weighted OD samples for each Municipality"

# Load the municipalities 
municipalities_ghs <- sf::st_read(connection,
                                  Id(schema="agile_gis_2025_rs",
                                     table="municipalities_ghs"))
# Select size of sample population
n_samples <- 200

# Create a list before looping for each municipality to generate sampling points
sampled_points_list_200 <- list()
library(tidyverse)
for (idx in 1:nrow(municipalities_ghs)){
  shapename <- municipalities_ghs$shapename[idx]
  municipality_geom <- municipalities_ghs$wkb_geometry[idx]
  n_od_samples <- n_samples
  masked_ghs_aoi <- mask(build_masked, vect(municipality_geom))
  
od <- spatSample(masked_ghs_aoi, size = n_od_samples, method = "weights", as.points = TRUE) 
 od_sf <- st_as_sf(od) |> st_transform(4326)
  
  # df with the point id and muncipality (shapename)
  od_sf <- od_sf |> 
    mutate(
      id_sample = row_number(),
      shapename = shapename
    ) |> 
    select(id_sample, shapename, geometry = geometry)
  
  # list that stores all the points
  sampled_points_list_200[[idx]] <- od_sf
}
# Create a dataframe from the previous list created during the loop
sampled_points_df_200 <- bind_rows(sampled_points_list_200)
# Export from R to PostgreSQL
DBI::dbWriteTable(connection, Id(schema="agile_gis_2025_rs","od_200_each"),sampled_points_df_200)
# time: 1468 ms
```

The following table shows the first 5 observations from the table _od_200_each_ where the municipality is indicated in the field shapename.

```{r}
#| echo: false
#| message: false
library(kableExtra)
kbl(slice(od_200_each,1:5))
```

### Snap weighted sampling points
#### For the Core Metropolitan Area

The sampling points of the OD matrix are not necessary located on the road network, therefore it is necessary to snap these points. Previous to snap the points, we made the largest component of the road network routable. (i.e. _ghs_ao_net_largest_). Subsequently, we indexed the geometries and ID variables to improve computational efficiency. We calculated the Euclidean distance between each sampling point and the nodes from the road network nodes. By limiting the results by one, we selected the closest node for each sampling point.

```{sql}
#| eval: false
#| echo: true
#| code-summary: "[ACR-3] How to snap the sampling points from the OD matrix to the road network"

--- (optional) Create "id" column as a pre-requisite for snapping if it is not created before.
ALTER TABLE od_300_v2
ADD COLUMN id INTEGER PRIMARY KEY GENERATED ALWAYS AS IDENTITY;

--- Similar to [ORN-3], we make the largest component routable
---   First we cast the variables according to pgrouting requirements
ALTER TABLE ghs_aoi_net_largest 
    ALTER COLUMN "toId" type bigint,
    ALTER COLUMN "fromId" type bigint,
    ALTER COLUMN "id" type bigint;
--- Second we create the auxiliary table ghs_aoi_net_largest_vertices_pgr
SELECT pgr_createverticesTable('ghs_aoi_net_largest',
the_geom:='geom',
source:='fromId',
target:='toId');

--- Create  indexes to improve queries performance.
CREATE INDEX idx_od_300_v2_id ON od_300_v2 USING btree(id);
CREATE INDEX idx_od_300_v2_geom ON od_300_v2 USING gist(geometry);
CREATE INDEX idx_ghs_aoi_net_largest_vertices_pgr ON ghs_aoi_net_largest_vertices_pgr USING btree(fid);
--- Snap the OD to the nodes of ghs_aoi_net_largest_vertices_pgr
EXPLAIN ANALYZE
CREATE TABLE od_300_snapped_v2 AS
SELECT DISTINCT ON (net.id)
  pt.id AS pt_id, --- original position of the Origin-Destination
  net.id AS snapped_id, --- snapped position of the Origin-Destination
  net.the_geom
FROM
(SELECT * FROM od_300_v2 AS pt) as pt
CROSS JOIN
LATERAL (SELECT
            *
        FROM
          ghs_aoi_net_largest_vertices_pgr AS net
        ORDER BY net.the_geom <->pt.geometry
        LIMIT 1) AS net;
---time: 35.474 ms
```

#### For the Intracity connectivity

The OD matrix is limited to each municipality instead of the core metropolitan area.  We created a numeric column named shapename_id to iterate each municipalities' networks.

```{sql}
#| eval: false
#| echo: true
#| code-summary: "[ACR-4-1] How to snap the sampling points from the OD matrix to each municipality road network"

--- Create a ID for each municipality
CREATE TABLE od_200_each_group_id AS
SELECT 
	*,
dense_rank() OVER (ORDER BY shapename) 
FROM od_200_each;
---time: 13.858 ms


```

First, we create an empty table that contained the final columns. 

```{r}
#| echo: true
#| eval: false
#| code-summary: "[ACR-4-2] How to snap the sampling points from the OD matrix to each municipality road network"
library(glue)
library(DBI)
od_200_each_group_id <- st_read(connection,"od_200_each_group_id")

table_snapped <- glue_sql("CREATE TABLE table_snapped 
                (pt_id numeric,
                snapped_id numeric,
                shapename_id varchar(200),
                shapename varchar(200),
                geom geometry)", .con =connection)
table_snapped_db <- dbGetQuery(conn=connection, table_snapped )

```

Then, we snapped the sampling points to the closest vertice of the network. The table _net_largest_1_vertices_pgr_ was created in the loop shown in code chunk [ORN-6]. The final table table_snapped contained the sampling point weighted on the building density for each of the municipalities. 

```{r}
#| echo: true
#| eval: false

## od_200_each_group_id
for (idx in unique(od_200_each_group_id$dense_rank)){
table_name_vertices <- glue("net_largest_{as.character(idx)}_vertices_pgr")
table_name_result <- glue("od_200_each_snapped_{as.character(idx)}")

dbExecute(glue_sql("DROP TABLE IF EXISTS {`table_name_result`}", .con=connection),
          conn=connection)  
query_1_snap <- glue_sql("CREATE TABLE {`table_name_result`} AS
SELECT DISTINCT ON (net.id)
       pt.id_sample AS pt_id,
       net.id AS snapped_id,
       pt.dense_rank AS shapename_id,
       pt.shapename,
       net.the_geom
FROM 
(SELECT * 
FROM 
    od_200_each_group_id
    WHERE
    dense_rank = {as.character(idx)}) as pt
CROSS JOIN
LATERAL (SELECT
        * 
        FROM {`table_name_vertices`} AS net
         ORDER BY net.the_geom <-> pt.geometry 
        LIMIT 1) AS net;", .con= connection)
result_1_snap <- dbGetQuery(conn=connection, query_1_snap)
 insert_query <- glue_sql("INSERT INTO table_snapped (pt_id, snapped_id, shapename_id,shapename, geom)
                            SELECT pt_id, snapped_id, shapename_id,shapename, the_geom 
                            FROM {`table_name_result`}", .con = connection)
  dbExecute(conn = connection, insert_query)
}

```


```{r}
#| eval: false
#| echo: false

for (idx in 1:9) {
  table_name_result <- glue("od_200_each_snapped_{idx}")
  
  insert_query <- glue_sql("
    INSERT INTO table_snapped (pt_id, snapped_id, shapename_id, shapename, geom)
    SELECT pt_id, snapped_id, shapename_id, shapename, the_geom
    FROM {`table_name_result`};
  ", .con = connection)
  
  dbExecute(conn = connection, insert_query)
}

```


**Troubleshooting 9** The query did not find the required tables. Since this iteration required the tables net_largest_x, where x is one of the municipalities id, we observed that these tables were in the public schema. We solved this by defining the search_path or the schema.  




## Regular OD

### Select Point of Interest (PoI)

```{sql}
#| eval: false
#| echo: true
#| code-summary: Hospitals

CREATE TABLE hospitals_poi AS
--- Add ICU bed capacity from ETLCNES to the hospitals
WITH all_hospitals AS (SELECT DISTINCT ON (cd_cnes)
	cd_cnes,
    ds_cnes,
    wkb_geometry AS geom,
		"QTLEITP1",
	"QTLEITP2",
	"QTLEITP3",
	"NAT_JUR",
	tp_risco,
	nm_razao_s,
	nm_municip
FROM
	hospitals AS geom_hospital
LEFT JOIN
	etlcnes_hospital_selected AS df_hospital
ON
	df_hospital."CNES"=geom_hospital.cd_cnes),
--- Define Point Of Interest (POI) using Area of Interest (AoI)
	 poi_hospitals AS (
SELECT 
    h.*
FROM 
    all_hospitals AS h,
    core_metropolitan_area AS poi
WHERE st_intersects(h.geom, poi.geom))
--- Add the original geometry and the snapped position
SELECT DISTINCT ON (h.cd_cnes)
	h.*,
    f.the_geom <-> h.geom AS distance,
    h.geom AS geom_hospital,
    f.the_geom AS geom_node
FROM poi_hospitals h
LEFT JOIN LATERAL
(SELECT 
    id, 
    the_geom
FROM ghs_aoi_net_largest_vertices_pgr AS net
ORDER BY
    net.the_geom <-> h.geom
LIMIT 1) AS f ON true

```


### Create Isochrones from PoI

Error: Openrouteservice API request failed
[3004] Parameter 'locations=22' is out of range. Maximum possible value is 5.

```{r}
#| eval: false
#| echo: true


library(sf)
library(openrouteservice)

# Remember to forget your API Rick, written by Rick.
openrouteservice::ors_api_key("5b3ce3597851110001cf6248ba03ca736c924d61aadb844a004da6fe")

hospitals_poi <- sf::st_read(connection, DBI::Id(schema="agile_gis_2025_rs" ,"hospitals_poi"))
library(tidyverse)
## Part I: PoI (1-5) 
isochrone_part_1 <- 
  ors_isochrones(st_coordinates(hospitals_poi$geom_node[1:5]),
                                range = 600,
                                interval = 600,
                                output = "sf") |>
                            mutate(ds_cnes = hospitals_poi$ds_cnes[1:5],
                                   cd_cnes = hospitals_poi$cd_cnes[1:5],
                                   id = hospitals_poi$id[1:5])
## Part II: PoI (6-10)
isochrone_part_2 <- 
  ors_isochrones(st_coordinates(hospitals_poi$geom_node[6:10]),
                                range = 600,
                                interval = 600,
                                output = "sf") |>
                            mutate(ds_cnes = hospitals_poi$ds_cnes[6:10],
                                   cd_cnes = hospitals_poi$cd_cnes[6:10],
                                   id = hospitals_poi$id[6:10])
## Part III: PoI (11-15)
isochrone_part_3 <- 
  ors_isochrones(st_coordinates(hospitals_poi$geom_node[11:15]),
                                range = 600,
                                interval = 600,
                                output = "sf") |>
                            mutate(ds_cnes = hospitals_poi$ds_cnes[11:15],
                                   cd_cnes = hospitals_poi$cd_cnes[11:15],
                                   id = hospitals_poi$id[11:15])
## Part IV: PoI (16-20)
isochrone_part_4 <- 
  ors_isochrones(st_coordinates(hospitals_poi$geom_node[16:20]),
                                range = 600,
                                interval = 600,
                                output = "sf") |>
                            mutate(ds_cnes = hospitals_poi$ds_cnes[16:20],
                                   cd_cnes = hospitals_poi$cd_cnes[16:20],
                                   id = hospitals_poi$id[16:20])
## Part V: PoI (20-22)
isochrone_part_5 <- 
  ors_isochrones(st_coordinates(hospitals_poi$geom_node[21:22]),
                                range = 600,
                                interval = 600,
                                output = "sf") |>
                            mutate(ds_cnes = hospitals_poi$ds_cnes[21:22],
                                   cd_cnes = hospitals_poi$cd_cnes[21:22],
                                   id = hospitals_poi$id[21:22])
## Merge 5 parts

isochrones_poi_hospitals <- rbind(isochrone_part_1,
                                  isochrone_part_2,
                                  isochrone_part_3,
                                  isochrone_part_4,
                                  isochrone_part_5 )  |>
                                select(-center) # the variable "center" is a list. Writting will return  no applicable method for st_write.
st_write(obj = isochrones_poi_hospitals,
         dsn = connection,
         Id(schema="agile_gis_2025_rs",
            table = "isochrones_poi_hospitals"))
```


### Create regular sampling points

FIrst he fcuntion: guck mal wo war diese Post 

```{sql}
#| eval: false
#| echo: true
#| code-summary: Regular OD

CREATE OR REPLACE FUNCTION I_Grid_Point_Series(geom geometry, x_side decimal, y_side decimal, spheroid boolean default false)
RETURNS SETOF geometry AS $BODY$
DECLARE
x_max decimal;
y_max decimal;
x_min decimal;
y_min decimal;
srid integer := 4326;
input_srid integer;
x_series DECIMAL;
y_series DECIMAL;
BEGIN
CASE st_srid(geom) WHEN 0 THEN
  geom := ST_SetSRID(geom, srid);
  RAISE NOTICE 'SRID Not Found.';
    ELSE
        RAISE NOTICE 'SRID Found.';
    END CASE;

    CASE spheroid WHEN false THEN
        RAISE NOTICE 'Spheroid False';
    else
        srid := 900913;
        RAISE NOTICE 'Spheroid True';
    END CASE;
    input_srid:=st_srid(geom);
    geom := st_transform(geom, srid);
    x_max := ST_XMax(geom);
    y_max := ST_YMax(geom);
    x_min := ST_XMin(geom);
    y_min := ST_YMin(geom);
    x_series := CEIL ( @( x_max - x_min ) / x_side);
    y_series := CEIL ( @( y_max - y_min ) / y_side );
RETURN QUERY
SELECT st_collect(st_setsrid(ST_MakePoint(x * x_side + x_min, y*y_side + y_min), srid)) FROM
generate_series(0, x_series) as x,
generate_series(0, y_series) as y
WHERE st_intersects(st_setsrid(ST_MakePoint(x*x_side + x_min, y*y_side + y_min), srid), geom);
END;
$BODY$ LANGUAGE plpgsql IMMUTABLE STRICT;

```

For the weig

```{sql}
#| eval: false
#| echo: true
#| code-summary: bbox

# Move core bbox 

CREATE TABLE core_metropolitan_bbox AS
SELECT 
	ST_Extent(geom)::geometry AS geom
FROM 
	core_metropolitan_area;

```


```{sql}
#| eval: false
#| echo: true
#| code-summary: Regular OD

# Move this to centrality hospitals

CREATE TABLE regular_point_od AS (
WITH multipoint_regular AS(
SELECT 
    I_Grid_Point_Series(geom, 0.037,0.037, false) AS geom
    FROM core_metropolitan_bbox AS geom),
point_regular AS(
SELECT 
    st_setsrid((st_dump(geom)).geom, 4326)::geometry(Point, 4326) AS geom
FROM  multipoint_regular)
SELECT 
    row_number() over() AS id,
    geom
FROM 
    point_regular);
```

For the Isochrones:

```{sql}
#| eval: false
#| echo: true

CREATE TABLE  regular_sampling_isochrones_ghs_aoi AS
WITH multipoint_regular AS(
SELECT 
    I_Grid_Point_Series(geometry, 0.00912,0.00912, false) AS geom,
    ds_cnes,
    cd_cnes
FROM 
	isochrones_poi_hospitals),
point_regular AS(
SELECT 
    st_setsrid((st_dump(geom)).geom, 4326)::geometry(Point, 4326) AS geom,
    ds_cnes,
    cd_cnes
FROM  multipoint_regular)
SELECT 
	cd_cnes,
    ds_cnes,
    geom,
	row_number() over() AS id
FROM 
    point_regular; 

```

### Snap regular sampling points

Regular

```{sql}
#| eval: false
#| echo: true

CREATE TABLE regular_sampling_isochrones_snapped_ghs_aoi AS
SELECT DISTINCT ON (net.id)
   	net.id,
   	ds_cnes,
   	cd_cnes,
   	net.the_geom  AS geom
FROM regular_sampling_isochrones_ghs_aoi AS pt
CROSS JOIN LATERAL 
	(SELECT 
		*
	FROM  ghs_aoi_net_largest_vertices_pgr AS net
	ORDER BY
	net.the_geom <-> pt.geom
	LIMIT 1) AS net;

```








