# Origin-Destination matrix

```{r}
#| eval: false
#| include: false
#| echo: false


# Snapping
od_200_each <- st_read(connection, Id(schema="agile_gis_2025_rs", table="od_200_each" ))
od_200_each_group_id <- st_read(connection, Id(schema="public", table="od_200_each_group_id"))  |> 
                          mutate(dense_rank = as.character(dense_rank)) 
```

Work in progress, waiting for acceptance. A

## Weighted OD

### Create weighted sampling points
#### For the Core Metropolitan Area

```{r}
#| eval: false
#| echo: true

# Import data
library(sf) # handle vector files
library(terra) # handle raster files
buildings <- terra::rast("/home/ricardors/agile-gscience-2024-rs-flood/data/source_data/GHS_BUILT_V_E2020_GLOBE_R2023A_4326_100_V1_0_RioGrandeDoSul.tif")
pop_ghs <- 300 # set up number of sampling points
## Select area of interest from the database
municipalities_ghs <- sf::read_sf(connection,
                                  DBI::Id(schema="agile_gis_2025_rs",
                                          table="municipalities_ghs"))
municipalities_ghs_reproj <- municipalities_ghs |>
                                            st_transform(crs(buildings))
build_cropped <- crop(buildings,
                      municipalities_ghs_reproj)
## Reproject for sampling
build_4326 <- terra::project(build_cropped,
                             crs(municipalities_ghs))
build_masked <- terra::mask(build_4326,
                            municipalities_ghs)
## Weighted sampling
set.seed(123)
od <- spatSample(build_masked,
                     pop_ghs,
                    "weights",
                     as.points = TRUE) |>
                st_as_sf() |>
                st_transform(4326)
## Rename column
names(od)[1] <- "building"
## Upload the sampling points to the database
sf::dbWriteTable(connection, DBI::Id(schema = "agile_gis_2025_rs",
table = "od_300_v2"), od)
```

#### For the Intracity connectivity

```{r}
#| eval: false
#| echo: true


# Load the municipalities 
municipalities_ghs <- sf::st_read(connection,
                                  Id(schema="agile_gis_2025_rs",
                                     table="municipalities_ghs"))
# Select size of sample population
n_samples <- 200

# Create a list before looping for each municipality to generate sampling points
sampled_points_list_200 <- list()
library(tidyverse)
for (idx in 1:nrow(municipalities_ghs)){
  shapename <- municipalities_ghs$shapename[idx]
  municipality_geom <- municipalities_ghs$wkb_geometry[idx]
  n_od_samples <- n_samples
  masked_ghs_aoi <- mask(build_masked, vect(municipality_geom))
  
od <- spatSample(masked_ghs_aoi, size = n_od_samples, method = "weights", as.points = TRUE) 
 od_sf <- st_as_sf(od) |> st_transform(4326)
  
  # df with the point id and muncipality (shapename)
  od_sf <- od_sf |> 
    mutate(
      id_sample = row_number(),
      shapename = shapename
    ) |> 
    select(id_sample, shapename, geometry = geometry)
  
  # list that stores all the points
  sampled_points_list_200[[idx]] <- od_sf
}
# Create a dataframe from the previous list created during the loop
sampled_points_df_200 <- bind_rows(sampled_points_list_200)
# Export from R to PostgreSQL
DBI::dbWriteTable(connection, Id(schema="agile_gis_2025_rs","od_200_each"),sampled_points_df_200)
```

### Snap weighted sampling points

#### For the Core Metropolitan Area

```{sql}
#| eval: false
#| echo: true
#| code-summary: snapping

# Create "id" column as a pre-requisite for snapping
ALTER TABLE od_300_v2
ADD COLUMN id INTEGER PRIMARY KEY GENERATED ALWAYS AS IDENTITY;

# Split this comments into several chunks
ALTER TABLE ghs_aoi_net_largest 
    ALTER COLUMN "toId" type bigint,
    ALTER COLUMN "fromId" type bigint,
    ALTER COLUMN "id" type bigint;

# Similar to chunk X
SELECT pgr_createverticesTable('ghs_aoi_net_largest',
the_geom:='geom',
source:='fromId',
target:='toId');

CREATE TABLE od_300_snapped_v2 AS
SELECT DISTINCT ON (net.id)
  pt.id AS pt_id, --- original position of the Origin-Destination
  net.id AS snapped_id, --- snapped position of the Origin-Destination
  net.geometry
FROM
(SELECT * FROM od_300_v2 AS pt) as pt
CROSS JOIN
LATERAL (SELECT
            *
        FROM
          ghs_aoi_net_largest_vertices_pgr AS net
        ORDER BY net.geom <->pt.geometry
        LIMIT 1) AS net;

```

#### For the Intracity connectivity

We created a numeric column named shapename_id to iterate for each of the municipalities' networks.

```{sql}
#| eval: false
#| echo: true


CREATE TABLE od_200_each_group_id AS
SELECT 
	*,
dense_rank() OVER (ORDER BY shapename) 
FROM od_200_each;

```

First, we create an empty table that contained the final columns. 

```{r}
#| echo: true
#| eval: false
library(glue)
library(DBI)
od_200_each_group_id <- st_read(connection,"od_200_each_group_id")

table_snapped <- glue_sql("CREATE TABLE table_snapped 
                (pt_id numeric,
                snapped_id numeric,
                shapename_id varchar(200),
                shapename varchar(200),
                geom geometry)", .con =connection)
table_snapped_db <- dbGetQuery(conn=connection, table_snapped )

```

We snapped the sampling points to the closest vertice of the network. The table net_largest_1_vertices_pgr was created in the loop shown in code chunk Y. The final table table_snapped contained the sampling point weighted on the building density for each of the municipalities. Indicar aquí también que se puso en el public.

```{r}
#| echo: true
#| eval: false

## od_200_each_group_id
for (idx in unique(od_200_each_group_id$dense_rank)){
table_name_vertices <- glue("net_largest_{as.character(idx)}_vertices_pgr")
table_name_result <- glue("od_200_each_snapped_{as.character(idx)}")

dbExecute(glue_sql("DROP TABLE IF EXISTS {`table_name_result`}", .con=connection),
          conn=connection)  
query_1_snap <- glue_sql("CREATE TABLE {`table_name_result`} AS
SELECT DISTINCT ON (net.id)
       pt.id_sample AS pt_id,
       net.id AS snapped_id,
       pt.dense_rank AS shapename_id,
       pt.shapename,
       net.the_geom
FROM 
(SELECT * 
FROM 
    od_200_each_group_id
    WHERE
    dense_rank = {as.character(idx)}) as pt
CROSS JOIN
LATERAL (SELECT
        * 
        FROM {`table_name_vertices`} AS net
         ORDER BY net.the_geom <-> pt.geometry 
        LIMIT 1) AS net;", .con= connection)
result_1_snap <- dbGetQuery(conn=connection, query_1_snap)
 insert_query <- glue_sql("INSERT INTO table_snapped (pt_id, snapped_id, shapename_id,shapename, geom)
                            SELECT pt_id, snapped_id, shapename_id,shapename, the_geom 
                            FROM {`table_name_result`}", .con = connection)
  dbExecute(conn = connection, insert_query)
}

```


Parche, falló lo de la tabla. Por si acaso,

```{r}
#| eval: false
#| echo: false

for (idx in 1:9) {
  table_name_result <- glue("od_200_each_snapped_{idx}")
  
  insert_query <- glue_sql("
    INSERT INTO table_snapped (pt_id, snapped_id, shapename_id, shapename, geom)
    SELECT pt_id, snapped_id, shapename_id, shapename, the_geom
    FROM {`table_name_result`};
  ", .con = connection)
  
  dbExecute(conn = connection, insert_query)
}

```


### Visualize Weigted OD

Add you removed OpenBuilding for QGIS.

```{=html}
<iframe width= "780" height="500" src="data/media/weighted_od.html" title "Weighted OD"></iframe>
```


```{r}
#| eval: false
#| echo: true

library(sf)
library(DBI)

## Import Data from PostgreSQL database
municipalities_ghs <- st_read(connection,
        DBI::Id(schema="agile_gis_2025_rs",
                table="municipalities_ghs"))
core_metropolitan_area <- st_read(connection,
        DBI::Id(schema="agile_gis_2025_rs",
                table="core_metropolitan_area"))        
od_300_snapped_v2 <- st_read(
                              connection,
                              DBI::Id(schema="agile_gis_2025_rs",
                                      table="od_300_snapped_v2"))
## Create interactive map
### styles
library(raster)
pal <- colorNumeric(c("#faebdd", "#cd1c4e", "#110c24"), values(build_masked),
  na.color = "transparent")
pal_fct <- colorFactor(
  palette = c('#e41a1c', '#377eb8', '#4daf4a', '#984ea3', '#ff7f00','#ffff33','#a65628', '#f781bf','#999999' ),
  domain = od_200_each$shapename
)
### map
library(leaflet)
leaflet() |>
  addTiles()  |>
  addProviderTiles("CartoDB.Positron", group = "CartoDB.Positron")  |>
  addProviderTiles("Esri.WorldImagery", group = "Esri.WorldImagery")  |>                    addPolygons(data=municipalities_ghs,
                    fillColor = "yellow",
                    fillOpacity = 0.05,
                    weight = 1.5,
                    color = "red",
                    dashArray = "1",
                    group= "Municipalities") |> 
  addPolygons(data=core_metropolitan_area,
                    fillColor = "yellow",
                    fillOpacity = 0.05,
                    weight = 2,
                    color = "red",
                    dashArray = "1",
                    group= "Core Metropolitan Area") |> 
  addCircleMarkers(data=od_300_snapped_v2,
                    fillColor = "yellow",
                    fillOpacity = 0.8,
                    weight = 1.5,
                    radius= 3,
                    color = "black",
                    dashArray = "1",
                    group= "Weighted OD - C.Metropolitan Area") |> 
   addCircleMarkers(data=od_200_each,
                    color = ~pal_fct(shapename),
                    fillOpacity = 0.8,
                    weight = 1.5,
                    radius= 3,
                    dashArray = "1",
                    group= "Weighted OD - Municipality") |> 
  addRasterImage(build_masked,
                 opacity=0.7,
                 colors = pal,
                 group="Bulding density") |> 
     addLegend(pal = pal, values = values(build_masked),
    title = "Building Density (m³)") |> 
 addLayersControl(
    baseGroups = c("CartoDB.Positron", "Esri.WorldImagery"),
    overlayGroups = c("Municipalities","Core Metropolitan Area", "Weighted OD - C.Metropolitan Area","Weighted OD - Municipality","Bulding density"),
    options = layersControlOptions(collapsed = FALSE)
  ) |>
  hideGroup(c("Municipalities","Core Metropolitan Area","Weighted OD - Municipality")) |>
  setView( lng =-51.21566 , lat = -30.04176, zoom = 12 ) 
```

## Regular OD

### Select Point of Interest (PoI)

```{sql}
#| eval: false
#| echo: true
#| code-summary: Hospitals

CREATE TABLE hospitals_poi AS
--- Add ICU bed capacity from ETLCNES to the hospitals
WITH all_hospitals AS (SELECT DISTINCT ON (cd_cnes)
	cd_cnes,
    ds_cnes,
    wkb_geometry AS geom,
		"QTLEITP1",
	"QTLEITP2",
	"QTLEITP3",
	"NAT_JUR",
	tp_risco,
	nm_razao_s,
	nm_municip
FROM
	hospitals AS geom_hospital
LEFT JOIN
	etlcnes_hospital_selected AS df_hospital
ON
	df_hospital."CNES"=geom_hospital.cd_cnes),
--- Define Point Of Interest (POI) using Area of Interest (AoI)
	 poi_hospitals AS (
SELECT 
    h.*
FROM 
    all_hospitals AS h,
    core_metropolitan_area AS poi
WHERE st_intersects(h.geom, poi.geom))
--- Add the original geometry and the snapped position
SELECT DISTINCT ON (h.cd_cnes)
	h.*,
    f.the_geom <-> h.geom AS distance,
    h.geom AS geom_hospital,
    f.the_geom AS geom_node
FROM poi_hospitals h
LEFT JOIN LATERAL
(SELECT 
    id, 
    the_geom
FROM ghs_aoi_net_largest_vertices_pgr AS net
ORDER BY
    net.the_geom <-> h.geom
LIMIT 1) AS f ON true

```


### Create Isochrones from PoI

Error: Openrouteservice API request failed
[3004] Parameter 'locations=22' is out of range. Maximum possible value is 5.

```{r}
#| eval: false
#| echo: true


library(sf)
library(openrouteservice)

# Remember to forget your API Rick, written by Rick.
openrouteservice::ors_api_key("5b3ce3597851110001cf6248ba03ca736c924d61aadb844a004da6fe")

hospitals_poi <- sf::st_read(connection, DBI::Id(schema="agile_gis_2025_rs" ,"hospitals_poi"))
library(tidyverse)
## Part I: PoI (1-5) 
isochrone_part_1 <- 
  ors_isochrones(st_coordinates(hospitals_poi$geom_node[1:5]),
                                range = 600,
                                interval = 600,
                                output = "sf") |>
                            mutate(ds_cnes = hospitals_poi$ds_cnes[1:5],
                                   cd_cnes = hospitals_poi$cd_cnes[1:5],
                                   id = hospitals_poi$id[1:5])
## Part II: PoI (6-10)
isochrone_part_2 <- 
  ors_isochrones(st_coordinates(hospitals_poi$geom_node[6:10]),
                                range = 600,
                                interval = 600,
                                output = "sf") |>
                            mutate(ds_cnes = hospitals_poi$ds_cnes[6:10],
                                   cd_cnes = hospitals_poi$cd_cnes[6:10],
                                   id = hospitals_poi$id[6:10])
## Part III: PoI (11-15)
isochrone_part_3 <- 
  ors_isochrones(st_coordinates(hospitals_poi$geom_node[11:15]),
                                range = 600,
                                interval = 600,
                                output = "sf") |>
                            mutate(ds_cnes = hospitals_poi$ds_cnes[11:15],
                                   cd_cnes = hospitals_poi$cd_cnes[11:15],
                                   id = hospitals_poi$id[11:15])
## Part IV: PoI (16-20)
isochrone_part_4 <- 
  ors_isochrones(st_coordinates(hospitals_poi$geom_node[16:20]),
                                range = 600,
                                interval = 600,
                                output = "sf") |>
                            mutate(ds_cnes = hospitals_poi$ds_cnes[16:20],
                                   cd_cnes = hospitals_poi$cd_cnes[16:20],
                                   id = hospitals_poi$id[16:20])
## Part V: PoI (20-22)
isochrone_part_5 <- 
  ors_isochrones(st_coordinates(hospitals_poi$geom_node[21:22]),
                                range = 600,
                                interval = 600,
                                output = "sf") |>
                            mutate(ds_cnes = hospitals_poi$ds_cnes[21:22],
                                   cd_cnes = hospitals_poi$cd_cnes[21:22],
                                   id = hospitals_poi$id[21:22])
## Merge 5 parts

isochrones_poi_hospitals <- rbind(isochrone_part_1,
                                  isochrone_part_2,
                                  isochrone_part_3,
                                  isochrone_part_4,
                                  isochrone_part_5 )  |>
                                select(-center) # the variable "center" is a list. Writting will return  no applicable method for st_write.
st_write(obj = isochrones_poi_hospitals,
         dsn = connection,
         Id(schema="agile_gis_2025_rs",
            table = "isochrones_poi_hospitals"))
```


### Create regular sampling points

FIrst he fcuntion: guck mal wo war diese Post 

```{sql}
#| eval: false
#| echo: true
#| code-summary: Regular OD

CREATE OR REPLACE FUNCTION I_Grid_Point_Series(geom geometry, x_side decimal, y_side decimal, spheroid boolean default false)
RETURNS SETOF geometry AS $BODY$
DECLARE
x_max decimal;
y_max decimal;
x_min decimal;
y_min decimal;
srid integer := 4326;
input_srid integer;
x_series DECIMAL;
y_series DECIMAL;
BEGIN
CASE st_srid(geom) WHEN 0 THEN
  geom := ST_SetSRID(geom, srid);
  RAISE NOTICE 'SRID Not Found.';
    ELSE
        RAISE NOTICE 'SRID Found.';
    END CASE;

    CASE spheroid WHEN false THEN
        RAISE NOTICE 'Spheroid False';
    else
        srid := 900913;
        RAISE NOTICE 'Spheroid True';
    END CASE;
    input_srid:=st_srid(geom);
    geom := st_transform(geom, srid);
    x_max := ST_XMax(geom);
    y_max := ST_YMax(geom);
    x_min := ST_XMin(geom);
    y_min := ST_YMin(geom);
    x_series := CEIL ( @( x_max - x_min ) / x_side);
    y_series := CEIL ( @( y_max - y_min ) / y_side );
RETURN QUERY
SELECT st_collect(st_setsrid(ST_MakePoint(x * x_side + x_min, y*y_side + y_min), srid)) FROM
generate_series(0, x_series) as x,
generate_series(0, y_series) as y
WHERE st_intersects(st_setsrid(ST_MakePoint(x*x_side + x_min, y*y_side + y_min), srid), geom);
END;
$BODY$ LANGUAGE plpgsql IMMUTABLE STRICT;

```

For the weig

```{sql}
#| eval: false
#| echo: true
#| code-summary: bbox

# Move core bbox 

CREATE TABLE core_metropolitan_bbox AS
SELECT 
	ST_Extent(geom)::geometry AS geom
FROM 
	core_metropolitan_area;

```


```{sql}
#| eval: false
#| echo: true
#| code-summary: Regular OD

# Move this to centrality hospitals

CREATE TABLE regular_point_od AS (
WITH multipoint_regular AS(
SELECT 
    I_Grid_Point_Series(geom, 0.037,0.037, false) AS geom
    FROM core_metropolitan_bbox AS geom),
point_regular AS(
SELECT 
    st_setsrid((st_dump(geom)).geom, 4326)::geometry(Point, 4326) AS geom
FROM  multipoint_regular)
SELECT 
    row_number() over() AS id,
    geom
FROM 
    point_regular);
```

For the Isochrones:

```{sql}
#| eval: false
#| echo: true

CREATE TABLE  regular_sampling_isochrones_ghs_aoi AS
WITH multipoint_regular AS(
SELECT 
    I_Grid_Point_Series(geometry, 0.00912,0.00912, false) AS geom,
    ds_cnes,
    cd_cnes
FROM 
	isochrones_poi_hospitals),
point_regular AS(
SELECT 
    st_setsrid((st_dump(geom)).geom, 4326)::geometry(Point, 4326) AS geom,
    ds_cnes,
    cd_cnes
FROM  multipoint_regular)
SELECT 
	cd_cnes,
    ds_cnes,
    geom,
	row_number() over() AS id
FROM 
    point_regular; 

```

### Snap regular sampling points

Regular

```{sql}
#| eval: false
#| echo: true

CREATE TABLE regular_sampling_isochrones_snapped_ghs_aoi AS
SELECT DISTINCT ON (net.id)
   	net.id,
   	ds_cnes,
   	cd_cnes,
   	net.the_geom  AS geom
FROM regular_sampling_isochrones_ghs_aoi AS pt
CROSS JOIN LATERAL 
	(SELECT 
		*
	FROM  ghs_aoi_net_largest_vertices_pgr AS net
	ORDER BY
	net.the_geom <-> pt.geom
	LIMIT 1) AS net;

```

### Visualize Regular OD

Add drop geometry porque tiene dos columnas, cuidadito

```{=html}
<iframe width="780" height=500 src="data/media/regular_od.html" title="Regular OD"></iframe>
```


```{r}
#| echo: true
#| eval: false

## Import data:
regular_sampling_isochrones_snapped_ghs_aoi <- 
  sf::st_read(connection, Id(schema="agile_gis_2025_rs",table="regular_sampling_isochrones_snapped_ghs_aoi"))

## style:
factpal <- colorFactor(topo.colors(22), isochrones_poi_hospitals$ds_cnes)
## Map
leaflet() |>
  addTiles()  |>
  addProviderTiles("CartoDB.Positron", group = "CartoDB.Positron")  |>
  addProviderTiles("Esri.WorldImagery", group = "Esri.WorldImagery")  |>   
  addProviderTiles("CartoDB.DarkMatter", group = "CartoDB.DarkMatter") |> 
  addCircleMarkers(data=hospitals_poi,
                    fillColor = "yellow",
                    fillOpacity = 0.6,
                    weight = 3,
                    radius = 5,
                    color = "red",
                    dashArray = "1",
                    group= "Hospitals PoI") |> 
  addPolygons(data=isochrones_poi_hospitals,
                    fillOpacity = 0.2,
                    weight = 2,
                    color = ~factpal(ds_cnes),
                    dashArray = "1",
                    group= "Isochrones") |> 
  addCircleMarkers(data=regular_sampling_isochrones_snapped_ghs_aoi,
                    fillOpacity = 0.8,
                    weight = 1.5,
                    radius= 3,
                    color =~factpal(ds_cnes),
                    dashArray = "1",
                    group= "Regular OD") |> 
 addLayersControl(
    baseGroups = c("CartoDB.Positron", "Esri.WorldImagery","CartoDB.DarkMatter"),
    overlayGroups = c("Hospitals PoI","Isochrones", "Regular OD"),
    options = layersControlOptions(collapsed = FALSE)
  ) |> 
    setView( lng =-51.01134 , lat = -29.93259, zoom = 11.5 ) 
```






